---
title: "Fitbit Case Study"
output: html_notebook
---
 
##Installing required packages

```{r}
install.packages('tidyverse')
install.packages('ggplot2')
install.packages('dplyr')
install.packages('lubridate')
install.packages('tidyr')
install.packages('corrplot')
install.packages('janitor')
```


##Installing required libraries

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyr)
library(corrplot)
library(janitor)
```


##Loading and reading dataset

```{r}
daily_activity <- read.csv('C:\\Users\\Owner\\OneDrive - Sify Technologies Limited\\Desktop\\Dataset\\Fitabase Data 4.12.16-5.12.16\\dailyActivity_merged.csv')
View(daily_activity)


daily_sleep = read.csv('C:\\Users\\Owner\\OneDrive - Sify Technologies Limited\\Desktop\\Dataset\\Fitabase Data 4.12.16-5.12.16\\sleepDay_merged.csv')

```


```{r}
n_distinct(daily_activity$Id)
n_distinct(daily_sleep$Id)
```

##Looking for missing values

```{r}
sum(is.na(daily_activity))
sum(is.na(daily_sleep))
```


##Checking for duplicatte values

```{r}
sum(duplicated(daily_activity))
sum(duplicated(daily_sleep))
```

##Checking the information 

```{r}
str(daily_activity)
glimpse(daily_activity)
glimpse(daily_sleep)
```

##Changing the date variable to date time format

```{r}

daily_activity$Date <- mdy(daily_activity$ActivityDate)
daily_sleep$Date <- mdy_hms(daily_sleep$SleepDay)
```


##Creating weekday

```{r}
daily_activity <- daily_activity %>% 
                  mutate(week_days = wday(Date,  
                                          label = T))
```

##Creating new variable active distance


```{r}
daily_activity = daily_activity %>%
   mutate(Active_distance = case_when(TotalSteps < 5000 ~ 'Sedentary',
                                   TotalSteps < 7499 ~ 'Less Active',
                                   TotalSteps < 9999 ~ 'Moderately                                                                Active',
                                   TotalSteps <12500 ~ 'Active',
                                   TotalSteps >= 12500 ~ 'Highly Active'))
```


##Creating new variable sleep time

```{r}
daily_sleep = daily_sleep %>%
  mutate(sleep_time = case_when(TotalMinutesAsleep < 420 ~ 'Not Enough',
                                TotalMinutesAsleep < 540 ~  'Good Sleep',
                                TotalMinutesAsleep > 540 ~ 'Over Sleep'))
```


##Creating new variable time in bed

```{r}
daily_sleep <- daily_sleep %>% 
  select(Id,
         Date,
         TotalSleepRecords, 
         TotalMinutesAsleep, 
         TotalTimeInBed) %>% 
  mutate(time_in_bed = (TotalTimeInBed-TotalMinutesAsleep))
```


##Merging two datasets

```{r}
final_data = merge(daily_activity, 
                   daily_sleep, 
                   by = c('Id', 'Date'))

```


##Creating correlation map

```{r}
data_corr = cor(final_data[, c(4, 5, 12, 13, 14, 15, 16,20)])
corrplot(data_corr, method = 'color')
corrplot(data_corr, method = 'number')
```

##Visualizing distribution sedentary minutes for minutes

```{r}
ggplot(daily_activity,
       aes(x=week_days,
           y=SedentaryMinutes, 
           fill = week_days))+
geom_boxplot()+ 
labs(title = 'Sedentary Level Distribution', 
     x='Day of the week',
     y = 'Sedentary (mins)')
```


##Visualizing active minutes for weekdays

```{r}
daily_activity <- daily_activity %>% 
                       mutate(Total_activity_mins =                                  VeryActiveMinutes+FairlyActiveMinutes+LightlyActiveMinutes )

ggplot(daily_activity, 
       aes(x = week_days, 
           y = Total_activity_mins, 
           fill = week_days)) +
geom_boxplot()+ 
labs(title = 'Active Minutes', 
     x = 'Days of the Week', 
     y = 'Total Active Minutes')
```


##Visualizing Active distance for weekdays

```{r}
ggplot(daily_activity, 
       aes(x = week_days, 
           y = TotalSteps, 
           fill = week_days)) +
geom_boxplot()+ 
labs(title = 'Active Distance', 
     x = 'Days of the Week', 
     y = 'Total Active Steps')

```


##Sleep distriibution for weekday

```{r}
ggplot(final_data, 
       aes(x = week_days, 
           y = TotalMinutesAsleep, 
           fill = week_days)) +
geom_boxplot()+ 
labs(title = 'Sleep Distribution for Week', 
     x = 'Days of the Week', 
     y = 'Total Sleep Time')
```


##Visualizing average total distance for everyday

```{r}
grouptime <- final_data %>% 
               group_by(Date) %>%
               summarise(mean_total = mean(TotalDistance))

ggplot(data = grouptime, 
       aes(x=Date, 
           y=mean_total))+
geom_bar(stat='identity',
         fill="lightblue")+ 
geom_line(color='red')+
theme(axis.text.x = element_text(angle=90))+
labs(title="Average Total Distance VS Date")  

```


##Visualizing Corelation between calories and total step

```{r}
ggplot(daily_activity, 
       aes(x = TotalSteps, 
           y = Calories))+
geom_point()+
geom_smooth(method = lm, 
            formula = y~x)+
labs(title='Daily Calories vs Total Steps')
```


##Visualizing correlation between calories and active minutes

```{r}
ggplot(daily_activity, 
       aes(x = VeryActiveMinutes, 
           y = Calories))+
geom_point()+
geom_smooth(method = lm, 
            formula = y~x)+
labs(title='Daily Calories vs Very Active Minutes')
```


##Visulizing correlation between Active minutes and total time sleep

```{r}
ggplot(final_data, 
       aes(x = VeryActiveMinutes, 
           y = TotalMinutesAsleep))+
geom_point()+
geom_smooth(method = lm, 
            formula = y~x)+
labs(title='Very Active Minutes vs Total Sleep Time')
```


##Visualizing sleep behaviour with sleep time

```{r}
ggplot(final_data_merged, 
       aes(x=sleep_time,
           y=TotalMinutesAsleep, 
           fill= sleep_time))+ 
geom_boxplot()+
labs(title= "Sleeping Behaviour", 
     x= "Sleep Type", 
     y= "Total Minutes Asleep",
     legends= "sleep category" )+ 
scale_fill_discrete(name="Sleep Category", 
                    labels= c("Good Sleep", "Not Enough", "Over Sleep"))
```

##Visualizing Active behaviour with total step

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(final_data, 
       aes(x=Active_distance,
           y=TotalSteps, 
           fill= Active_distance))+ 
geom_boxplot()+
labs(title= "Active Behaviour", 
     x= "Active Type", 
     y= "Total Steps",
     legends= "Active category" )+ 
scale_fill_discrete(name="Active Category", 
                    labels= c("Active", "Very Active", "Less Active",                                  'Moderately Active', 'Sedentary'))


```

##Visualizing correlation between sedentary minutes and total sleep 

```{r}
ggplot(final_data, 
       aes(x = SedentaryMinutes, 
           y = TotalMinutesAsleep))+
geom_point()+
geom_smooth(method = lm, formula = y~x)+
labs(title='Sedentary Minutes vs Total Sleep Time')
```


##Creating new variable total active minutes

```{r}

final_data_merged <-   final_data_merged%>%                                                      mutate(TotalActiveMinutes = VeryActiveMinutes+
                                  FairlyActiveMinutes+
                                  LightlyActiveMinutes)


```


##Visualizing correlation between active minutes and sleep time

```{r}
ggplot(final_data_merged, 
       aes(x = TotalActiveMinutes, 
           y = TotalMinutesAsleep))+
geom_point()+
geom_smooth(method = lm, formula = y~x)+
labs(title='Total Active Minutes vs Total Sleep Time')
```

##Visualizing correlation between active minutes and total distance 

```{r}
ggplot(final_data, aes(x = TotalDistance, y = VeryActiveMinutes))+
  geom_point()+
  geom_smooth(method = lm, formula = y~x)+
  labs(title='Very Active Minutes vs Total Distance')
```


##Visualizing correlation between fairly active minutes and total distance

```{r}
ggplot(final_data, aes(x = FairlyActiveMinutes, y = TotalDistance))+
  geom_point()+
  geom_smooth(method = lm, formula = y~x)+
  labs(title='Fairly Active Minutes vs Total Distance')


```

